Do Alternate Wetting and Drying Irrigation Technology and Nitrogen Rates Affect Rice Sheath Blight?

### 2020-06-25 analysis/paper/paper.Rmd
Add keywords, key message, clean up figures, add references

While the effects of climate change directly on rice disease have been studied, the effects of mitigation practices have not.
Rice farmers may be hesitant to adopt water saving technologies because changing rice production practices could have unintended effects on the diseases that occur in the rice crop.
This is a reasonable concern because of the well documented effects of N [@Slaton2003; @Tang2007] and the effects of alternating wet and dry periods [@Castilla1996] on sheath blight.
In order to gain a better understanding of what the effects of adopting AWD on rice sheath blight may be, we established field experiments that incorporated irrigation technology and N fertilisation.

@@

Area under the disease progress stairs (AUDPS) [@simko2012area] was calculated for both disease incidence and severity using R, version 3.5.0 [@R2018], package `agricolae`, version 1.2-8, [@deMendiburu2017].
Due to the use of a severity scale for data collection, the severity data were converted to the midpoint percent value of the scale range for each value.

The correlations between tiller sheath sheath blight incidence and tiller and leaf sheath blight severity were tested using Spearman’s correlation test, `cor.test`, [@R2018].
No correlation was found, so the analysis was completed for each variable independently with no assumed interaction.

As most of the data's residuals did not meet assumptions for normality, the analysis was carried out using Bayesian multivariate generalised linear mixed models implemented in the R package `MCMCglmm` [@MCMCglmm2010], version 2.26.
Six models were created where the the dependent variables were 2015 tiller sheath blight incidence, 2015 tiller sheath blight severity, 2015 leaf sheath blight severity; 2016 tiller sheath blight incidence, 2016 tiller sheath blight severity; 2016 leaf sheath blight severity, respectively.
The independent variables were the irrigation management and N rate with replicate treated as a random variable.
The base levels (control treatments) for the analyses for 2015 were: N rate -- N0, irrigation management -- puddled and for 2016 they were: N rate -- N60, irrigation management -- puddled

@@

In both experiments we were unable to detect any statistically clear effect of AWD on sheath blight that led to increases in the disease that could hinder adoption of the technology.
In fact, in only one instance were we able to detect any statistically clear effect of AWD on sheath blight.
In the 2016 experiment the tiller sheath blight severity was clearly lower for the AWD treatment than puddled treatment (Fig.&nbsp;\@ref(fig:fig-3)d, \@ref(fig:fig-5)d), indicating a possible adverse effect of using AWD on tiller sheath blight severity under high sheath blight pressure.

@@

Climate change is predicted to decrease sheath blight risk in the Korean peninsula over the longer term.
However, over the next 10 -- 20 years the risk is predicted to remain the same as from 2000 -- 2010 [@Kim2015].
Therefore, it is likely to remain an important disease in the near future with little resistance available in current varieties.
However, based on these studies the use of AWD technologies would not appear to be a factor that increases the incidence or severity of rice sheath blight disease.

Readers may find a reproducible research compendium that recreates this paper, that includes all data generated from this project and code used for analysis as a reproducible research compendium from <https://doi.org/...> [@Sparks2019].

@@

@@

<div id="refs"></div>
\onecolumn

@@

##### Sheath blight progress for water management treatments

@@

##### Sheath blight progress for N treatments

@@

##### Sheath Blight Incidence Boxplot for water management

@@

##### Tiller sheath blight incidence boxplot for N rate
```{r fig-4, fig.cap="Sheath blight severity progress was measured in the 2015 at five assessment points and 2016 experiment at four points. Sheath blight incidence was rated as the number of infected tillers divided by the total number of tillers per hill and used to calculate the area under the disease progress stairs (AUDPS). Sheath blight severity was rated on an unevenly spaced categorical scale and converted to the mid--point percentage value to calculate the AUDPS. Five different N treatments were applied to split plots depending on the experiment. In the 2015 experiment three N rate treatments were applied: no N (N0), 100&nbsp;kg&nbsp;/&nbsp;ha N (N100) and 180&nbsp;kg&nbsp;/&nbsp;ha N (N120). In the 2016 experiment, two N rate treatments were applied: 60&nbsp;kg&nbsp;/&nbsp;ha N (N60) and 180&nbsp;kg&nbsp;/&nbsp;ha N (N180)."}

@@

##### Tiller Estimates Plots

@@

plot_tiller_estimates_paper(models[[5]])
mtext(text = "d", side = 3, line = 0, adj = 1, cex = 0.75)
```

##### Leaf estimates plots

@@

par(mfcol = c(1, 2))

plot_tiller_estimates_paper(models[[3]], main_title = "2015")
mtext(text = "Leaf Severity", side = 2, line = 6)
mtext(text = "a", side = 3, line = 0, adj = 1, cex = 0.75)

@@

##### pagebreak

### Colophon {-}

>>>>

While the effects of climate change directly on rice disease have been studied, the effects of mitigation practices have not.
Rice farmers may be hesitant to adopt water saving technologies because changing rice production practices could have unintended effects on the pests and diseases that occur in the rice crop.
This is a reasonable concern because of the well documented effects of N [@Slaton2003; @Tang2007] and the effects of alternating wet and dry periods [@Castilla1996] on sheath blight.
In order to gain a better understanding of what the effects of adopting AWD on rice sheath blight may be, we established field experiments that incorporated irrigation technology and N fertilisation.

@@

Area under the disease progress stairs (AUDPS) [@simko2012area] was calculated for both disease incidence and severity using R, version 4.0.0 [@R2020], package `agricolae`, version 1.3-3, [@agricolae2020].
Due to the use of a severity scale for data collection, the severity data were converted to the midpoint percent value of the scale range for each value.

The correlations between tiller sheath sheath blight incidence and tiller and leaf sheath blight severity were tested using Spearman’s correlation test, `cor.test`, [@R2020].
No correlation was found, so the analysis was completed for each variable independently with no assumed interaction.

As most of the data's residuals did not meet assumptions for normality, the analysis was carried out using Bayesian multivariate generalised linear mixed models implemented in the R package `MCMCglmm`, version 2.29, [@MCMCglmm2010].
Six models were created where the the dependent variables were 2015 tiller sheath blight incidence, 2015 tiller sheath blight severity, 2015 leaf sheath blight severity; 2016 tiller sheath blight incidence, 2016 tiller sheath blight severity; 2016 leaf sheath blight severity, respectively.
The independent variables were the irrigation management and N rate with replicate treated as a random variable.
The base levels (control treatments) for the analyses for 2015 were: N rate -- N0, irrigation management -- puddled and for 2016 they were: N rate -- N60, irrigation management -- puddled

@@

In both experiments we were unable to detect any statistically clear effect of AWD on sheath blight that led to increases in the disease that could hinder adoption of the technology.
In only one instance were we able to detect any statistically clear effect of AWD on sheath blight.
In the 2016 experiment the tiller sheath blight severity was clearly lower for the AWD treatment than puddled treatment (Fig.&nbsp;\@ref(fig:fig-3)d, \@ref(fig:fig-5)d), indicating a possible adverse effect of using AWD on tiller sheath blight severity under high sheath blight pressure.

@@

Diseases are the only issue that farmers face when growing rice.
Recently, Lorica *et al.* [@Lorica2020] have also found that AWD did not affect rodent damage in crops.
Findings such as these are useful as IRRI promotes water-saving technologies like AWD because farmers can be assured that the changes will not increase the risks from these pests and pathogens.

Climate change is predicted to decrease sheath blight risk in the Korean peninsula over the longer term [@Kim2015].
However, over the next 10 -- 20 years the risk is predicted to remain the same as from 2000 -- 2010 [@Kim2015].
Therefore, it is likely to remain an important disease in the near future with little resistance available in current varieties.
However, based on these studies the use of AWD technologies would not appear to be a factor that increases the incidence or severity of rice sheath blight disease.

@@

## Reproducibility Statement {-}

Readers may find a reproducible research compendium that recreates this paper, which includes all data generated from this project and code used for analysis as a reproducible research compendium from <https://doi.org/...> [@Sparks2019].

@@

<div id="refs"></div>

\onecolumn

@@

##### Sheath blight progress for water management treatments {-}
\\

@@

##### Sheath blight progress for N treatments {-}
\\

@@

##### Sheath Blight Incidence Boxplot for water management {-}
\\

@@

##### Tiller sheath blight incidence boxplot for N rate {-}
\\

```{r fig-4, fig.cap="Sheath blight severity progress was measured in the 2015 at five assessment points and 2016 experiment at four points. Sheath blight incidence was rated as the number of infected tillers divided by the total number of tillers per hill and used to calculate the area under the disease progress stairs (AUDPS). Sheath blight severity was rated on an unevenly spaced categorical scale and converted to the mid--point percentage value to calculate the AUDPS. Five different N treatments were applied to split plots depending on the experiment. In the 2015 experiment three N rate treatments were applied: no N (N0), 100&nbsp;kg&nbsp;/&nbsp;ha N (N100) and 180&nbsp;kg&nbsp;/&nbsp;ha N (N120). In the 2016 experiment, two N rate treatments were applied: 60&nbsp;kg&nbsp;/&nbsp;ha N (N60) and 180&nbsp;kg&nbsp;/&nbsp;ha N (N180)."}

@@

##### Tiller Estimates Plots {-}
\\

@@

plot_tiller_estimates_paper(models[[5]])
mtext(text = "d", side = 3, line = 0, adj = 1, cex = 0.75)
par(mfcol = c(1, 1))
```

##### Leaf estimates plots {-}
\\

@@

par(mfcol = c(2, 1))
plot_tiller_estimates_paper(models[[3]], main_title = "2015")
mtext(text = "Leaf Severity", side = 2, line = 6)
mtext(text = "a", side = 3, line = 0, adj = 1, cex = 0.75)

@@

<!-- ##### pagebreak -->

\newpage

# Colophon {-}
